---
title: "Alcala_STA445_HW5"
author: "Angelica Alcala"
date: "2023-11-09"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Question Notes:
Question 1: For the infmort data set, the country names are the rownames() of the data.frame. You will
need to make a new column within the infmort data.frame for the countries, so that you can use ggrepel
functions correctly.
Question 2: Please add to the graph the best fit line, and make the annotated information based on the
lm fit information (coefficients and R-squared), rather than the random values provided.
1

# Question 1

```{r, warning = FALSE}
library(faraway)
data(infmort)
library(ggplot2)
library(ggrepel)
```


## a

The rownames() of the table gives the country names and you should create a new column that contains the country names. *rownames

```{r}
infmort2 <- infmort %>%
  mutate(Country = row.names(.)) %>%
  select('Country', everything()) 
row.names(infmort2) = NULL
head(infmort2)
```


## b

Create scatter plots with the log10() transformation inside the aes() command.


```{r}
ggplot(infmort2, aes(x = log10(income), y = log10(mortality))) +
  geom_point(aes(color = oil)) +
  facet_wrap(.~ region)
```


## c

Create the scatter plots using the scale_x_log10() and scale_y_log10(). Set the major and minor breaks to be useful and aesthetically pleasing. Comment on which version you find easier to read.


To be honest I felt like the log10 function was easier to read, although I can that there are probably instances where the scale_y_log10 function might help visualize data better. For this instance I found log10 easier and allowled for better visualization of the data.


```{r}
ggplot(infmort2, aes(x = income, y = mortality)) +
  geom_point(aes(color = oil)) +
  facet_wrap(.~ region) +
  scale_y_log10(breaks=c(100,200,400,600,800),
                  minor=c(100,125,150,175,
                        seq( 1, 100,by=10 )))
```

## d.

The package ggrepel contains functions geom_text_repel() and geom_label_repel() that mimic the basic geom_text() and geom_label() functions in ggplot2, but work to make sure the labels don’t overlap. Select 10-15 countries to label and do so using the geom_text_repel() function.

```{r}
samp.infmort <- infmort2 %>%
  mutate(Country2 = ifelse(str_detect(Country, pattern = 'g'), Country, NA))
samp.infmort
```



```{r}
ggplot(samp.infmort, aes(x = log10(income), y = log10(mortality))) +
  geom_point(aes(color = oil)) +
  facet_wrap(.~ region) +
  geom_text_repel(aes(label = Country2)) + 
  labs(x=latex2exp::TeX('$\\log_{10}$(Income)'), y=latex2exp::TeX('$\\log_{10}$(Mortality)'),
       title = 'Mortality as a Function of Income')
```


# Question 2

```{r}
datasets::trees
```


## a. 

Create a regression model for y = Volume as a function of x = Height.

```{r}
reg.model <- lm(Volume ~ Height, data = trees)
reg.model
```


## b.

Using the summary command, get the y-intercept and slope of the regression line.

```{r}
summary(reg.model)
```


### The y intercept is - 87.124, the slope is 1.543


## c.

Using ggplot2, create a scatter plot of Volume vs Height.


## d. 

Create a nice white filled rectangle to add text information to using by adding the following annotation layer.

```{r}
annotate('rect', xmin=65, xmax=75, ymin=60, ymax=74,
         fill='white', color='black') #+
```


## e.

Add some annotation text to write the equation of the line $y_i=−8.2+15∗x_i$ in the text area.

See 5.3

## f. 
Add annotation to add $R^2 = 0.358$
 

## g.

Add the regression line in red. The most convenient layer function to uses is geom_abline(). It appears that the annotate doesn’t work with geom_abline() so you’ll have to call it directly.