---
title: 'STA 445 Exam #1'
author: "Angelica Alcala"
date: "`r format(Sys.time(),  '%B %d, %Y')`"
output: pdf_document
---

```{r, include=FALSE, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
### Here are suggested libraries, you may load others if you wish.
library(readxl)
library(tidyverse)
library(lubridate)
```



# Exam Questions

## Question 1 [30 points]

You will be asked to produce two functions. The first (`vecsummary`) will create a summary output for a given numerical vector.  The second (`mean_split`) will provide ABOVE/BELOW strings depending on if the element in the vector is above or below the mean of the vector.  Please answer each sub question below.

**a.** For some data to analyze with your functions, load the data set `cars` from base R. The `cars` set contains two variables, `speed` and `dist`, of which there are $50$ total observations.

```{r}
data(cars) 
mean(cars$speed)
mean(cars$dist)

```

**b.** Produce the function `vecsummary`. The function should accept as an input a numerical vector and return a data.frame with columns `Size`, `Min`, `Max`, `Min.Max`, `Mean`, in that order. The size here refers to the number of elements (`length`) in the vector. The element `Min.Max` is the ratio `Min / Max`. The function should return a data.frame with proper labels for each column, in the order specified above..

```{r}
vecsummary <- function(x){
  Size <- length(x)
  Min.Max <- Min/Max
  Min <- min(data.frame$x)
  Max <- max(data.frame$x)
  Mean <- mean(x)
output <- data.frame('Size','Min','Max','Min.Max','Mean')
return(output)
}
```

**c.** Demonstrate that the `vecsummary` function works properly by displaying the results of your function when analyzing the variable `speed` within the `cars` object.

```{r}
#vecsummary(cars$speed)
```

**d.** Produce the function `mean_split`. The function should accept as an input a numerical vector and return a vector of ABOVE/BELOW determined for each element being above or below the mean value of the vector.

```{r}
mean_split <- function(x){
  output <- ifelse(x < mean(x), 'BELOW','ABOVE')
  return(output)
}
```

**e.** Demonstrate that the `mean_split` function works properly by displaying the results of your function when analyzing the variable `dist` within the `cars` object.

```{r}
mean_split(cars$dist)
```

## Question 2 [20 points]

In the area marked as `???` write a regular expression that accomplishes the requested task. Use the `strings` object to demonstrate your answer is correct by giving a minimum of two TRUE and two FALSE responses.

**a.** An expression that can determine if a string contains any of the following words: `Captain`, `Chief`, `Colonel`, `Cadet`, `Cook`.

```{r, eval =TRUE}
#strings <- c("Captain", "Chief", "Colonel", "Cadet", "Cook", "hello","test", "Frank")
#data.frame(string = strings) %>%
  #mutate(result = str_detect(string, 'Captain', 'Chief', `Colonel`, `Cadet`, `Cook`) )
```

**b.** An expression that can determine if a string begins with 2 of any upper or lowercase letter followed by any digit 4 times.

```{r, eval=FALSE}
### Change to eval=TRUE or remove eval option
#strings <- c('a5555', 'hello')
#data.frame( string = strings ) %>%
  #mutate( result = str_detect(string, 'D', 'd', '{4}') )
```

**c.** An expression that can determine if a string begins with `bar` or ends in `foo`.

```{r, eval=FALSE}
### Change to eval=TRUE or remove eval option
strings <- c()
data.frame( string = strings ) %>%
  mutate( result = str_detect(string, '???') )
```

## Question 3 [15 points]

**a.** Traveling to `Europe/Rome`, you are told that your flight will arrive at 2:00 PM local time on June 15th, 2024.  Create the object `Rome.Arrival` for this time stamp, be sure to include the timezone. Display the object to screen.

```{r}
Rome.Arrival <- mdy_hm('Jun 15 2024, 2:00 PM', tz = 'Europe/Rome')
Rome.Arrival
```

**b.** You want to make a call back to the USA when you arrive.  What time will it be in `US/Arizona` when you arrive in Rome?  Return/display the results of an R calculation here.

```{r}
UStime <- Rome.Arrival %>%
  with_tz(tzone = 'US/Arizona')
UStime
```

**c.** You are set to return to the USA on August 10th, 2024 at 10:00 PM (Arizona).  Assuming all of these dates/times are not delayed, how many __days__ will you be traveling? (Return a calculated time object that displays days, hours, minutes, seconds).

```{r}
USreturn <- mdy_hm('Aug 10 2024, 10:00 PM', tz = 'US/Arizona')
USreturn
```

```{r}
Travel <- interval(Rome.Arrival, USreturn)
Traveltime <- as.period(Travel)
Traveltime
```



## Question 4 [25 points]

The `Exam1_Data_F23.xlsx` file contains data developed across multiple trips (`ID`) where information was collected for four different `Sites` at four different `Periods`. Two measurements were taken for each observations and were recorded as `X` and `Y`. There are potentially many observations of the same `Period` and `Site`. The data was collected across two trips that are available in the `Exam1_Data.xlsx` file on Sheets 2 and 3.

**a.** Read in the two different data sets on Sheets 2 and 3. Combine the data into a single data frame named `Q4data`. There should be 5 columns: `ID`, `Site`, `Period`, `X`, `Y`, totaling 100 observations. Demonstrate this by returning to screen the structure of the `Q4data` object. This can be done using the `str()` function. Do __not__ display any of the data.frames in your final answer, only the structure of the combined data frame (a tibble is acceptable).

```{r}
Sheet2 <- read_excel(path = 'Exam1_Data_F23.xlsx', sheet = 2, range = 'A1:E47')
Sheet3 <- read_excel(path = 'Exam1_Data_F23.xlsx', sheet = 3, range = 'A1:E55')
Q4data <- full_join(Sheet2,Sheet3)
str(Q4data)
```

**b.** We are interested in counting the number of `Site` and `Period` combinations.  Create a `wide` table that has the `Period` as columns, the `Site` as rows, and the counts for each pair within the table.

```{r}
Q4data2 <- Q4data %>%
  pivot_wider(Site, names_from = ID, values_from = Period)
Q4data2
```

**c.** Finally, we want to make a visual assessment of the total `X` and total `Y` observed across the four different observation sets `ID`.  Group the data by `Site` and `Period` and summarize the total `X` and total `Y` found for each combination.  Save this as a new data.frame `plotme` and display the `head`. *Hint: Finding the total requires we take the sum.*

```{r}

```

**d.** Use the data.frame `plotme` to view the total X against the total Y as a scatter graph, and color the scatter based on the `Site`. 

```{r}

```